A family of accessibility measures derived from spatial interaction principles

### 2025-02-10 paper/manuscript.qmd
Adjusted the simple numerical example

```{r load-packages, include=FALSE, cache=FALSE}
library(bibliometrix) # Comprehensive Science Mapping Analysis
library(dplyr) # A Grammar of Data Manipulation
library(flextable) # Functions for Tabular Reporting
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(ggraph) # An Implementation of Grammar of Graphics for Graphs and Networks
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(here) # A Simpler Way to Find Your Files
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(tidytext) # Text Mining using 'dplyr', 'ggplot2', and Other Tidy Tools
# library(huxtable) # Easily Create and Style Tables for LaTeX, HTML and Other Formats
# library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
# library(patchwork) # The Composer of Plots
# library(RColorBrewer) # ColorBrewer Palettes
# library(sf) # Simple Features for R
# library(scales) # Scale Functions for Visualization
# library(shadowtext) # Shadow Text Grob and Layer
# library(skimr) # Compact and Flexible Summaries of Data
# library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
# library(tidyr) # Tidy Messy Data
# library(tmap) # Thematic Maps
# library(ggpmisc) # Miscellaneous Extensions to 'ggplot2'
# library(ggrepel) # Automatically Position Non-Overlapping Text Labels with 'ggplot2'
# library(ggspatial) # Spatial Data Framework for ggplot2
# library(gridExtra) # Miscellaneous Functions for "Grid" Graphics 
# library(here) # A Simpler Way to Find Your Files
```

@@

The following is the known accessibility matrix. The $V_{ij}$, the number of jobs spatially accessible at $i$ from destination $j$. The TOTALs are the $V_i$ and $V_j$, the total number of jobs spatially available at $i$ and the number of jobs spatially available from $j$.
```{r}
known_matrix <- matrix(c(95,23,42, 27,378,45, 38,36,106), ncol = 3, byrow = TRUE)
#unknowns <- matrix(c(rep(NA,9)), ncol=3, byrow=TRUE)
P_i <- known_matrix |> rowSums()
O_j <- known_matrix |> colSums()
V <- ifelse(sum(P_i)==sum(O_j), sum(O_j),NA)
knowns_longer <- cbind(known_matrix,P_i)
knowns_longer <- rbind(knowns_longer,cbind(t(O_j),T))

rownames(knowns_longer) <- c("1","2","3", "TOTAL") 
colnames(knowns_longer) <- c("1","2","3", "TOTAL")

knowns_longer #|> as.data.frame() |> round() |> gt(rownames_to_stub=TRUE)
```

And the matrix of cost of travel from $i$ (rows) to $j$ (columns) is as follows:
```{r}
C <- matrix(c(2, 15, 5, 15, 2, 10, 5, 10, 2), ncol = 3, byrow = TRUE)
C_longer <-C
rownames(C_longer) <- c("1","2","3") 
colnames(C_longer) <- c("1","2","3") 
C_longer #|> as.data.frame() |> gt(rownames_to_stub=TRUE)
```

@@

In solving the example, each origin (row) all the opportunities at each destination are available. Likewise, each destination (column) offers all opportunities to each origin: 
```{r}
O_P_non <- t(matrix(rep(O_j, 3), 3, 3))
O_P_non_longer <- O_P_non
rownames(O_P_non_longer) <- c("1","2","3") 
colnames(O_P_non_longer) <- c("1","2","3")
O_P_non_longer
```

The resulting matrix of $S_{ij}$ along with the $S_i$ and $S_j$ totals for this example:
```{r}
S_non <- O_P_non / C

P_i_non <- S_non |> rowSums()
O_j_non <- S_non |> colSums()
S_TOTAL_non <- ifelse(sum(P_i_non|>round())==sum(O_j_non|>round()), sum(S_non), NA)
S_non_longer <- cbind(S_non,P_i_non)
S_non_longer <- rbind(S_non_longer, cbind(t(O_j_non),S_TOTAL_non))

rownames(S_non_longer) <- c("1","2","3", "TOTAL") 
colnames(S_non_longer) <- c("1","2","3", "TOTAL")

S_non_longer |> round(digits = 2) #|> as.data.frame()  |> gt(rownames_to_stub=TRUE)
```

@@

@@

@@

```{r}
K <- V/sum(S_non) #S_non -> (1/C)* t(matrix(rep(O_j, 3), 3, 3))
V_tot <- K * S_non

P_i_tot <- V_tot |> rowSums()
D_j_tot <- V_tot |> colSums()
V_TOTAL_tot <- ifelse(sum(P_i_tot)==sum(D_j_tot), sum(V_tot), NA)
V_tot_longer <- cbind(V_tot,P_i_tot)
V_tot_longer <- rbind(V_tot_longer, cbind(t(D_j_tot),V_TOTAL_tot))

rownames(V_tot_longer) <- c("1","2","3", "TOTAL") 
colnames(V_tot_longer) <- c("1","2","3", "TOTAL")

V_tot_longer |> round() #|> as.data.frame()  |> gt(rownames_to_stub=TRUE)
```

While we can see that the 'number of potentially reachable opportunities' `r V` results, the total access at a each $i$ or $j$ is not equal to the known known total access ends (because these constraints are not specified and thus not satisfied by the model). This approach is not yet seen in accessibility literature, but could be an intelligible way to compare accessibility values ($V_i$ or $V_j$) without considering the population matrix.

@@

Returning to our simple example, the opportunity-constrained singly-constrained case would yield the following $B_{j}$ values assuming $A_{i}$ is 1:

@@

```{r}
B_j <- 1/(P_i %*% (1/C))
B_j_longer <- B_j
rownames(B_j_longer) <- c("i..") 
colnames(B_j_longer) <- c("1","2","3") 
B_j_longer 
```

@@

$V_{ij}''^{(2)}$ values:
```{r}
V_dest <- outer(P_i, as.numeric(B_j) * O_j) * (1/C) #we're not doing standard matrix multiplication here. We use 'outer' for element-wise multiplication, then element-wise multiplication of the resulting 3*3 matrix with the 3*3 impedance table

P_i_dest <- V_dest |> rowSums()
O_j_dest <- V_dest |> colSums()

V_TOTAL_dest <- ifelse(all.equal(sum(P_i_dest),sum(O_j_dest)), sum(V_dest), NA)
V_dest_longer <- cbind(V_dest,P_i_dest)
V_dest_longer <- rbind(V_dest_longer, cbind(t(O_j_dest),V_TOTAL_dest))

rownames(V_dest_longer) <- c("1","2","3", "TOTAL") 
colnames(V_dest_longer) <- c("1","2","3", "TOTAL")

V_dest_longer |> round() #|> as.data.frame()  |> gt(rownames_to_stub=TRUE)
```

@@

In $V_i''^{(1)}$, the following is $A_i$ assuming $B_{j}$ is 1:

@@

```{r}
A_i <- 1/((1/C) %*% O_j)
A_i_longer <- A_i
rownames(A_i_longer) <- c("1","2","3") 
colnames(A_i_longer) <- c("j...") 
A_i_longer 
```

(AGAIN NOTE: when $A_i$ is multiplied by $O_j*f(c_{ij})$, it is a proportion (like spatial availability's total balancing factor) representing how much of $P_i$ is allocated from each $i$, I think. e.g. each $V_{ij}$ is the proportional opportunity&travel-cost balancing factor of $A_{i} W_j^{(2)}f(c_{ij})$ multiplied by $P_i$)

The resulting $V_{ij}''^{(1)}$ values:
```{r}
V_orig <- outer(as.numeric(A_i) * P_i, O_j) * (1/C) #we're not doing standard matrix multiplication here. We use 'outer' for element-wise multiplication, then element-wise multiplication of the resulting 3*3 matrix with the 3*3 impedance table

P_i_orig <- V_orig |> rowSums()
O_j_orig <- V_orig |> colSums()

V_TOTAL_orig <- ifelse(all.equal(sum(P_i_orig),sum(O_j_orig)), sum(V_orig), NA)
V_orig_longer <- cbind(V_orig,P_i_orig)
V_orig_longer <- rbind(V_orig_longer, cbind(t(O_j_orig),V_TOTAL_orig))

V_orig_longer |> round() #|> as.data.frame()  |> gt(rownames_to_stub=TRUE)
```

@@

## Doubly-constrained case (population-opportunity constrained)

In this case, $V_{ij}$ would be:
```{r}
doubly_constrained_gravity_model <- function(
    #CREDIT: I converted the python code from (https://github.com/SadraDaneshvar/Gravity_Model) to R
  O,  # Origin production weights vector
  D,  # Destination attraction weights vector
  impedance_matrix,  # the resulting travel cost impedance matrix
  error_threshold = 0.001,  # Error threshold for stopping condition
  improvement_threshold = 1e-6  # Improvement threshold for stopping condition
) {

@@

    # Calculate Tij
    Tij <- outer(Ai * O, Bj_new * D) * impedance_matrix #Tij <- outer(Ai * O, Bj_new * D) * impedance_matrix -->

@@

The $A_i$ and $B_j$ matrices:
```{r}
A_i_doubly <- dc_results$Ai
B_j_doubly <- dc_results$Bj

@@

The trip matrix (notice, all summed up opportunities and populations are equivalent to our knowns)
```{r}
V_doubly <- dc_results$Tij

@@

Nice.

@@

## Checking singly-constrained case against spatial availability
```{r}
sp_avail <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

@@

```{r}

@@

```{r}
A_ij <- matrix(O_j) %*% (1/(t((matrix(O_j))) %*% (1/C))) |> t()
A_ij_longer <- A_ij
rownames(A_ij_longer) <- c("1","2","3") 
colnames(A_ij_longer) <- c("1","2","3") 
A_ij_longer 
```

Starting off with this opportunity matrix, namely - at i=1, only a fixed number of opportunities are accessible to this origin. same for i=2, i=3.
```{r}
(matrix(rep(O_j, 3), 3, 3))
```

$V_{ij}$ values:
```{r}
V_orig <- (A_ij) * t(S_non) #S_non -> (1/C)* t(matrix(rep(O_j, 3), 3, 3)) #if you transpose the matrix here, that's another way of generating the answer.

>>>>

```{r load-packages, include=FALSE, cache=FALSE}
library(bibliometrix) # Comprehensive Science Mapping Analysis
library(dplyr) # A Grammar of Data Manipulation
library(flextable) # Functions for Tabular Reporting
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(ggraph) # An Implementation of Grammar of Graphics for Graphs and Networks
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(here) # A Simpler Way to Find Your Files
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(tidytext) # Text Mining using 'dplyr', 'ggplot2', and Other Tidy Tools
```

@@

The following is the known accessibility matrix. The $V_{ij}$, the number of jobs spatially accessible at $i$ from destination $j$. The TOTALs are the $V_i$ and $V_j$, the total number of jobs spatially available at $i$ and the number of jobs spatially available from $j$.

The following vectors are the population by origin ($O_i$) and the opportunities by destination $D_j$:

<!--Check here, the number of physicians per 10,000 in Canada in 2022 was 24.97
https://en.wikipedia.org/wiki/List_of_countries_by_number_of_physicians -->
```{r}
#| label: simple-numerical-example-opportunities-and-population

id <- c("1", "2", "3")
O_i <- c(4, 10, 6) 
D_j <- c(160, 150, 180)
LU <- data.frame(id, O_i, D_j)
```

```{r}
LU
```

Think of $O_i$ as population in 10,000 and $D_j$ as number of physicians. The Provider-to-Population-Ratio (PPR) is:
```{r}
sum(LU$D_j) / sum(LU$O_i)
```

And the cost of travel from $i$ to $j$ is as follows:
```{r}
#| label: simple-numerical-example-cost

# Simulate a table with OD cost for the simple example, beginning with the ids, the cost, and then two impedance functions 
C <- expand.grid(oid = c("1", "2", "3"),
                 did = c("1", "2", "3")) |>
  mutate(cost = c(2, 15, 5, 15, 2, 10, 5, 10, 2)) |>
  mutate(f1 = exp(-1.00000 * cost^2), 
         f2 = exp(-0.01000 * cost^2),
         f3 = exp(-0.00001 * cost^2))
```

```{r}
C |> select(oid, did, cost)
```

For comparison, use three different impedance functions:
$$
\begin{array}{l}
f_1(c_{ij}) = exp(-c_{ij}^2)\\
f_2(c_{ij}) = exp(-0.01\cdot c_{ij}^2)\\
f_3(c_{ij}) = exp(-0.00001\cdot c_{ij}^2)
\end{array}
$$

@@

In solving the example, each origin (row) all the opportunities at each destination are available. Likewise, each destination (column) offers all opportunities to each origin: 
```{r}
#| label: simple-numerical-example-OD-table

OD <- C |>
  left_join(LU |>
              select(-D_j),
            by = c("oid" = "id")) |>
  left_join(LU |>
              select(-O_i),
            by = c("did" = "id"))
```

```{r}
unc_acc_ij <- OD |>
  group_by(oid) |>
  reframe(V_unc_ij_1 = D_j * f1,
            V_unc_ij_2 = D_j * f2,
            V_unc_ij_3 = D_j * f3)

unc_acc <- unc_acc_ij |>
  group_by(oid) |>
  summarize(V_unc_i_1 = sum(V_unc_ij_1),
            V_unc_i_2 = sum(V_unc_ij_2),
            V_unc_i_3 = sum(V_unc_ij_3))

unc_acc
```

Compare the accessibility differences and between pairs of zone using the two different impedance functions: 
```{r}
unc_acc$V_unc_i_1[2] - unc_acc$V_unc_i_1[1]
unc_acc$V_unc_i_2[2] - unc_acc$V_unc_i_2[1]
unc_acc$V_unc_i_3[2] - unc_acc$V_unc_i_3[1]

unc_acc$V_unc_i_1[2]/unc_acc$V_unc_i_1[1]
unc_acc$V_unc_i_2[2]/unc_acc$V_unc_i_2[1]
unc_acc$V_unc_i_3[2]/unc_acc$V_unc_i_3[1]
```

This illustrates how the differences and ratios are not meaningful.

The total accessibility compared to the number of opportunities:
```{r}
unc_acc <- unc_acc |> 
  left_join(LU |>
              select(-D_j),
            by = c("oid" = "id"))

unc_acc  |>
  summarize(V_unc_1 = sum(V_unc_i_1),
            V_unc_2 = sum(V_unc_i_2),
            V_unc_3 = sum(V_unc_i_3))

LU |>
  summarize(O = sum(D_j))
```

And the ratio of accessibility to population compared to the ratio of opportunities to population:
```{r}
# Accessibility divided by population
unc_acc |> 
  mutate(v_unc_1 = V_unc_i_1/O_i,
            v_unc_2 = V_unc_i_2/O_i,
            v_unc_3 = V_unc_i_3/O_i)

# Sum of the accessibility divided by the total population
unc_acc |> 
  summarize(v_unc_1 = sum(V_unc_i_1)/sum(O_i),
            v_unc_2 = sum(V_unc_i_2)/sum(O_i),
            v_unc_3 = sum(V_unc_i_3)/sum(O_i))

# Sum of the opportunities divided by the total population
LU |> 
  summarize(PPR = sum(D_j)/sum(O_i))
```

Notice the wild differences. Accessibility according to impedance function 3 basically says that everyone has access to all opportunities all at once. Accessibility according to function 1, on the other hand, says that each population has accessibility to around 3 opportunities, which is absurd: the number of opportunities available locally are in the order of 150-180. One should not have to adjust the impedance function to obtain "reasonable" results [EXAMPLE]; instead, the impedance function should reflect the travel behavior.

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

@@

**END OF PREVIOUS TEXT**

@@

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

@@

**END OF PREVIOUS TEXT**

Calculate the balancing factors/proportionality constants:
```{r}
k_tot <- unc_acc_ij |> 
  summarize(k1 = sum(LU$D_j)/sum(V_unc_ij_1),
            k2 = sum(LU$D_j)/sum(V_unc_ij_2),
            k3 = sum(LU$D_j)/sum(V_unc_ij_3))
```

Use these proportionality constants to calculate the total opportunity constrained accessibility:
```{r}
tot_acc_ij <- unc_acc_ij |>
  mutate(V_tot_ij_1 = k_tot$k1 * V_unc_ij_1,
         V_tot_ij_2 = k_tot$k2 * V_unc_ij_2,
         V_tot_ij_3 = k_tot$k3 * V_unc_ij_3)

tot_acc <- tot_acc_ij |>
  group_by(oid) |>
  summarize(V_tot_i_1 = sum(V_tot_ij_1),
            V_tot_i_2 = sum(V_tot_ij_2),
            V_tot_i_3 = sum(V_tot_ij_3))

tot_acc |>
  select(oid, starts_with("V_"))
```

Compare the accessibility differences and between pairs of zone using the two different impedance functions: 
```{r}
tot_acc$V_tot_i_1[2] - tot_acc$V_tot_i_1[1]
tot_acc$V_tot_i_2[2] - tot_acc$V_tot_i_2[1]
tot_acc$V_tot_i_3[2] - tot_acc$V_tot_i_3[1]

tot_acc$V_tot_i_1[2]/tot_acc$V_tot_i_1[1]
tot_acc$V_tot_i_2[2]/tot_acc$V_tot_i_2[1]
tot_acc$V_tot_i_3[2]/tot_acc$V_tot_i_3[1]
```

This illustrates how the differences and ratios are now meaningful.

The total accessibility compared to the number of opportunities:
```{r}
tot_acc <- tot_acc |> 
  left_join(LU |>
              select(-D_j),
            by = c("oid" = "id"))

tot_acc  |>
  summarize(V_tot_1 = sum(V_tot_i_1),
            V_tot_2 = sum(V_tot_i_2),
            V_tot_3 = sum(V_tot_i_3))

LU |>
  summarize(O = sum(D_j))
```

This shows how the number of opportunities is preserved.

And the ratio of accessibility to population compared to the ratio of opportunities to population:
```{r}
# Accessibility divided by population
tot_acc |> 
  mutate(v_tot_1 = V_tot_i_1/O_i,
            v_tot_2 = V_tot_i_2/O_i,
            v_tot_3 = V_tot_i_3/O_i)

# Sum of the accessibility divided by the total population
tot_acc |> 
  summarize(v_tot_1 = sum(V_tot_i_1)/sum(O_i),
            v_tot_2 = sum(V_tot_i_2)/sum(O_i),
            v_tot_3 = sum(V_tot_i_3)/sum(O_i))

# Sum of the opportunities divided by the total population
LU |> 
  summarize(PPR = sum(D_j)/sum(O_i))
```

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

While we can see that the 'number of potentially reachable opportunities' V results, the total access at a each $i$ or $j$ is not equal to the known known total access ends (because these constraints are not specified and thus not satisfied by the model). This approach is not yet seen in accessibility literature, but could be an intelligible way to compare accessibility values ($V_i$ or $V_j$) without considering the population matrix.

@@

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

Returning to our simple example, the attraction-constrained case would yield the following $B_{j}$ values assuming $A_{i}$ is 1:

@@

**END OF PREVIOUS TEXT**

Balancing factors $B_j$:
```{r}
B_j <- OD |> 
  group_by(did) |>
  summarize(B_j_1 = 1/sum(O_i * f1),
            B_j_2 = 1/sum(O_i * f2),
            B_j_3 = 1/sum(O_i * f3))

B_j
```

Use these proportionality constants to calculate the total opportunity constrained accessibility:
```{r}
opp_acc_ij <- OD |>
  left_join(B_j,
            by = "did") |>
  mutate(V_opp_ij_1 = B_j_1  * O_i * D_j * f1,
         V_opp_ij_2 = B_j_2  * O_i * D_j * f2,
         V_opp_ij_3 = B_j_3  * O_i * D_j * f3)

opp_acc <- opp_acc_ij |>
  group_by(oid) |>
  summarize(V_opp_i_1 = sum(V_opp_ij_1),
            V_opp_i_2 = sum(V_opp_ij_2),
            V_opp_i_3 = sum(V_opp_ij_3))

opp_acc  |>
  select(oid, starts_with("V_"))
```

Compare the accessibility differences and between pairs of zone using the two different impedance functions: 
```{r}
opp_acc$V_opp_i_1[2] - opp_acc$V_opp_i_1[1]
opp_acc$V_opp_i_2[2] - opp_acc$V_opp_i_2[1]
opp_acc$V_opp_i_3[2] - opp_acc$V_opp_i_3[1]

opp_acc$V_opp_i_1[2] / opp_acc$V_opp_i_1[1]
opp_acc$V_opp_i_2[2] / opp_acc$V_opp_i_2[1]
opp_acc$V_opp_i_3[2] / opp_acc$V_opp_i_3[1]
```

This illustrates how the differences and ratios are now meaningful.

The total accessibility compared to the number of opportunities:
```{r}
opp_acc <- opp_acc |> 
  left_join(LU |>
              select(-D_j),
            by = c("oid" = "id"))

opp_acc  |>
  summarize(V_opp_1 = sum(V_opp_i_1),
            V_opp_2 = sum(V_opp_i_2),
            V_opp_3 = sum(V_opp_i_3))

LU |>
  summarize(O = sum(D_j))
```

This shows how the number of opportunities is preserved.

And the ratio of accessibility to population compared to the ratio of opportunities to population:
```{r}
# Accessibility divided by population
opp_acc |> 
  mutate(v_opp_1 = V_opp_i_1/O_i,
            v_opp_2 = V_opp_i_2/O_i,
            v_opp_3 = V_opp_i_3/O_i)

# Sum of the accessibility divided by the total population
opp_acc |> 
  summarize(v_opp_1 = sum(V_opp_i_1)/sum(O_i),
            v_opp_2 = sum(V_opp_i_2)/sum(O_i),
            v_opp_3 = sum(V_opp_i_3)/sum(O_i))

# Sum of the opportunities divided by the total population
LU |> 
  summarize(PPR = sum(D_j)/sum(O_i))
```

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

@@

@@

**END OF PREVIOUS TEXT**

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

Returning to our simple example, the production-constrained case would yield the following $A_{j}$:

@@

**END OF PREVIOUS TEXT**

Balancing factors $A_i$:
```{r}
A_i <- OD |> 
  group_by(oid) |>
  summarize(A_i_1 = 1/sum(D_j * f1),
            A_i_2 = 1/sum(D_j * f2),
            A_i_3 = 1/sum(D_j * f3))

A_i
```

Use these proportionality constants to calculate the total opportunity constrained accessibility:
```{r}
pop_acc_ij <- OD |>
  left_join(A_i,
            by = "oid") |>
  mutate(V_pop_ij_1 = A_i_1  * O_i * D_j * f1,
         V_pop_ij_2 = A_i_2  * O_i * D_j * f2,
         V_pop_ij_3 = A_i_3  * O_i * D_j * f3)

pop_acc <- pop_acc_ij |>
  group_by(did) |>
  summarize(V_pop_i_1 = sum(V_pop_ij_1),
            V_pop_i_2 = sum(V_pop_ij_2),
            V_pop_i_3 = sum(V_pop_ij_3))

pop_acc  |>
  select(did, starts_with("V_"))
```

Compare the accessibility differences and between pairs of zone using the two different impedance functions: 
```{r}
pop_acc$V_pop_i_1[2] - pop_acc$V_pop_i_1[1]
pop_acc$V_pop_i_2[2] - pop_acc$V_pop_i_2[1]
pop_acc$V_pop_i_3[2] - pop_acc$V_pop_i_3[1]

pop_acc$V_pop_i_1[2] / pop_acc$V_pop_i_1[1]
pop_acc$V_pop_i_2[2] / pop_acc$V_pop_i_2[1]
pop_acc$V_pop_i_3[2] / pop_acc$V_pop_i_3[1]
```

This illustrates how the differences and ratios are now meaningful.

The total accessibility compared to the number of opportunities:
```{r}
pop_acc <- pop_acc |> 
  left_join(LU |>
              select(-O_i),
            by = c("did" = "id"))

pop_acc  |>
  summarize(V_pop_1 = sum(V_pop_i_1),
            V_pop_2 = sum(V_pop_i_2),
            V_pop_3 = sum(V_pop_i_3))

LU |>
  summarize(P = sum(O_i))
```

This shows how the population is preserved.

And the ratio of accessibility to population compared to the ratio of opportunities to population:
```{r}
# Accessibility divided by population
pop_acc |> 
  mutate(v_pop_1 = V_pop_i_1/O_i,
            v_pop_2 = V_pop_i_2/O_i,
            v_pop_3 = V_pop_i_3/O_i)

# Sum of the accessibility divided by the total population
pop_acc |> 
  summarize(v_pop_1 = sum(V_pop_i_1)/sum(D_j),
            v_pop_2 = sum(V_pop_i_2)/sum(D_j),
            v_pop_3 = sum(V_pop_i_3)/sum(D_j))

# Sum of the opportunities divided by the total population
LU |> 
  summarize(PPR = sum(O_i))/sum(D_j)
```

**PREVIOUS TEXT FOLLOWS: REVISE DISCUSSION**

@@

**END OF PREVIOUS TEXT**

## Doubly-constrained case (population-opportunity constrained)
NOTE: The example needs to be revised in this case to have the population and opportunities match, as in, one person-one opportunity

<!--
In this case, $V_{ij}$ would be:
```{r}
#| eval: false

doubly_constrained_gravity_model <- function(
    #CREDIT: I converted the python code from (https://github.com/SadraDaneshvar/Gravity_Model) to R
  O,  # Origin production weights vector
  D,  # Destination attraction weights vector
  impedance_matrix,  # the resulting travel cost impedance matrix
  error_threshold = 0.001,  # Error threshold for stopping condition
  improvement_threshold = 1e-6  # Improvement threshold for stopping condition
) {

@@

    # Calculate Tij
    Tij <- outer(Ai * O, Bj_new * D) * impedance_matrix #Tij <- outer(Ai * O, Bj_new * D) * impedance_matrix

@@

The $A_i$ and $B_j$ matrices:
```{r}
#| eval: false

A_i_doubly <- dc_results$Ai
B_j_doubly <- dc_results$Bj

@@

The trip matrix (notice, all summed up opportunities and populations are equivalent to our knowns)
```{r}
#| eval: false

V_doubly <- dc_results$Tij

@@

Nice. 
-->

@@

## Checking singly-constrained case against spatial availability
```{r}
#| eval: false

sp_avail <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

@@

```{r}
#| eval: false

@@

```{r}
#| eval: false

A_ij <- matrix(O_j) %*% (1/(t((matrix(O_j))) %*% (1/C))) |> t()
A_ij_longer <- A_ij
rownames(A_ij_longer) <- c("1","2","3") 
colnames(A_ij_longer) <- c("1","2","3") 
A_ij_longer 
```

Starting off with this opportunity matrix, namely - at i=1, only a fixed number of opportunities are accessible to this origin. same for i=2, i=3.
```{r}
#| eval: false

(matrix(rep(O_j, 3), 3, 3))
```

$V_{ij}$ values:
```{r}
#| eval: false

V_orig <- (A_ij) * t(S_non) #S_non -> (1/C)* t(matrix(rep(O_j, 3), 3, 3)) #if you transpose the matrix here, that's another way of generating the answer.
